Predictive Processing Model for an Autonomous Agent

### 2016-12-20 doc/paper.tex
using pointmass1d/3d as alternative system
preparing plots for section (currently) 5.1 example M1

\usepackage[T1]{fontenc}
%% % \usepackage[longnamesfirst,sort,nonamebreak]{natbib}
%% \usepackage[normalem]{ulem}
%% \usepackage[table]{xcolor}
\usepackage[utf8]{inputenc}
%% \usepackage{amsmath}
%% \usepackage{amssymb}
%% \usepackage{amstext}
%% \usepackage{array}
%% \usepackage{authblk}
%% \usepackage{caption}
%% \usepackage{color}
%% \usepackage{epsfig} % for postscript graphics files
%% \usepackage{fixltx2e}
%% \usepackage{float}
%% % \usepackage{fullpage}
%% \usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{hyperref}
%% \usepackage{latexsym}
%% \usepackage{lmodern}
%% \usepackage{longtable}
%% \usepackage{makeidx}  % allows for indexgeneration
%% \usepackage{marvosym}
%% \usepackage{multirow}
%% \usepackage{subcaption}
%% % \usepackage{subfigure}
%% \usepackage{textcomp}
%% \usepackage{titling}
%% \usepackage{wasysym}
%% \usepackage{wrapfig}

@@

\usepackage[style=authoryear,backend=biber,bibencoding=utf8,natbib]{biblatex}
\bibliography{actinf}
\addbibresource{actinf.bib}

@@

Long history, many models: control, development, RL, neuroscience, ...\ldots{}

@@

\subsection{General Models}

@@

The basic building block (BBB) contains an adaptive forward model
(predictor) whose output is the BBB's output. We assume direct
feedback from that module that the output is connected to. The BBB'
inputs are an explicit goal and the prediction error with respect to
the last prediction at ${t-1}$. These inputs taken together encode the
state. Thus the forward model should in principle be able to pick up
on different relations depending on the state. Graphical
representation is given in \autoref{fig:M1}.

@@

This modification of BBB M1 receives only the error as an input. The
model's graph is shown in \autoref{fig:M2}. This model will not be
able to distinguish between situations demanding different
predictions.

@@

@@

@@

>>>>

\usepackage[T1]{fontenc}
%% % \usepackage[longnamesfirst,sort,nonamebreak]{natbib}
%% \usepackage[normalem]{ulem}
%% \usepackage[table]{xcolor}
\usepackage[utf8]{inputenc}
%% \usepackage{amsmath}
%% \usepackage{amssymb}
%% \usepackage{amstext}
%% \usepackage{array}
%% \usepackage{authblk}
%% \usepackage{caption}
%% \usepackage{color}
%% \usepackage{epsfig} % for postscript graphics files
%% \usepackage{fixltx2e}
%% \usepackage{float}
%% % \usepackage{fullpage}
%% \usepackage{graphics} % for pdf, bitmapped graphics files
\usepackage{hyperref}
%% \usepackage{latexsym}
\usepackage{lmodern}
%% \usepackage{longtable}
%% \usepackage{makeidx}  % allows for indexgeneration
%% \usepackage{marvosym}
%% \usepackage{multirow}
%% \usepackage{subcaption}
\usepackage[tight,footnotesize]{subfigure}
%% \usepackage{textcomp}
%% \usepackage{titling}
%% \usepackage{wasysym}
%% \usepackage{wrapfig}

@@

\usepackage[style=authoryear,backend=biber,bibencoding=utf8,natbib]{biblatex}
\bibliography{actinf}
% \addbibresource{actinf.bib}

@@

Long history, many models: control, development, RL, neuroscience, prediction and prediction error (residual), ...\ldots{}

@@

\subsection{Basis building blocks}

@@

The basic building block (BBB) contains an adaptive forward model
(predictor) whose output is the BBB's output. We assume direct
feedback from the module that the output is connected to. The BBB'
inputs are an explicit goal and the prediction error with respect to
the last prediction at ${t-1}$. These inputs taken together encode the
state. Thus the forward model should in principle be able to pick up
on different relations depending on the state. Graphical
representation is given in \autoref{fig:M1}. The closed-loop dynamics
of this model are described in detail in \autoref{sec:exprs_basic_ex_m1}.

FIXME: Fundamental limitation: can only learn monotonic relationships so
far. Show to fail, indicate amendment that can solve non-monotonic
cases (by observing PE over time, by using reward modulation, ...).

@@

This modification of BBB M1 receives only the error as an input. The
model's graph is shown in \autoref{fig:M2}. This model will not be
able to distinguish between situations demanding different
predictions. FIXME: It can be shown to fail when such a situation is produced
by experiment design.

@@

\subsection{E2P}

A third building block is the intermodal map, specifically mapping
between an exteroceptive and the proprioceptive modality (thus
E2P). We propose to realize this slightly differently from M1 /
M2. The E2P module is bootstrapped by passively observing
corresponding exteroceptive and proprioceptive data and learning those
correspondences. Once bootstrapped, the E2P can be used to predict
possibly ambigous proprioceptive configurations from the exeroceptive
sensations they are believed to produce, meaning which have been
observed in the past to do so.

@@

Notes for experiments:

@@

\subsection{Basic behavioural example M1}
\label{sec:exprs_basic_ex_m1}

\begin{itemize}
\item Basic operation:
  \autoref{fig:basic_op_200_knn}, \autoref{fig:basic_op_200_soesgp},
  \autoref{fig:basic_op_1000_knn}, \autoref{fig:basic_op_1000_soesgp}
  show variations on the basic operation theme. We are evaluating a
  k-NN and a Gaussian Process based forward model each over 200 and
  1000 timesteps interaction on a one-dimensional identity system. The
  forward model starts off in a zero-state, the first prediction being
  zero. Any resulting prediction error drives an adaptation on the
  forward model output towards reduced error for a given goal, error
  input pair. We sweep this
  input space once at numsteps/2 and once at the end of the
  experiment which demonstrates the change in the overall model
  response.
  \item Compare identity system with small aberration system: now we
    can do the same for a 1D system while introducing aberrations in
    the system's response function.
  \item Failure under non-monotonicity: if the aberrations become
    systemically non-monotonic the learning process fails (FIXME: at least I
    predict that ;)
\item How PE is low when revisiting previously observed
  configurations: longterm PE plot of 10000 timesteps and averaged PE
  in addition to instantaneous PE
% \item operation using non-identity motor transfer function
\end{itemize}

\begin{figure}
  \begin{center}
  \subfigure[System sweep response]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_knn/fig_001_scattermatrix_reduced.pdf}%
    \label{fig:basic_op_200_knn_fig1}
  }
  \subfigure[Goal, prediction, error]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_knn/fig_003_aie_experiment_plot_basic.pdf}
    \label{fig:basic_op_200_knn_fig2}
  }
  \subfigure[Model sweep t = 100]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_knn/fig_002_colormeshmatrix_reduced.pdf}
    \label{fig:basic_op_200_knn_fig3}
  }
  \subfigure[Model sweep t = 200]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_knn/fig_004_colormeshmatrix_reduced.pdf}
    \label{fig:basic_op_200_knn_fig4}
  }
  \end{center}
  \caption{\label{fig:basic_op_200_knn}Basic operation 200 timesteps knn.}
\end{figure}

\begin{figure}
  \begin{center}
  \subfigure[System sweep response]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_soesgp/fig_001_scattermatrix_reduced.pdf}%
    \label{fig:basic_op_200_soesgp_fig1}
  }
  \subfigure[Goal, prediction, error]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_soesgp/fig_003_aie_experiment_plot_basic.pdf}
    \label{fig:basic_op_200_soesgp_fig2}
  }
  \subfigure[Model sweep t = 100]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_soesgp/fig_002_colormeshmatrix_reduced.pdf}
    \label{fig:basic_op_200_soesgp_fig3}
  }
  \subfigure[Model sweep t = 200]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_200steps_soesgp/fig_004_colormeshmatrix_reduced.pdf}
    \label{fig:basic_op_200_soesgp_fig4}
  }
  \end{center}
  \caption{\label{fig:basic_op_200_soesgp}Basic operation 200 timesteps GP.}

\end{figure}
\begin{figure}
  \begin{center}
  \subfigure[System sweep response]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_knn/fig_001_scattermatrix_reduced.pdf}%
    \label{fig:basic_op_1000_knn_fig1}
  }
  \subfigure[Goal, prediction, error]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_knn/fig_003_aie_experiment_plot_basic.pdf}
    \label{fig:basic_op_1000_knn_fig2}
  }
  \subfigure[Model sweep t = 500]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_knn/fig_002_colormeshmatrix_reduced.pdf}
    \label{fig:basic_op_1000_knn_fig3}
  }
  \subfigure[Model sweep t = 1000]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_knn/fig_004_colormeshmatrix_reduced.pdf}
    \label{fig:basic_op_1000_knn_fig4}
  }
  \end{center}
  \caption{\label{fig:basic_op_1000_knn}Basic operation 1000 timesteps knn.}
\end{figure}

\begin{figure}
  \begin{center}
  \subfigure[System sweep response]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_soesgp/fig_001_scattermatrix_reduced.pdf}%
    \label{fig:basic_op_1000_soesgp_fig1}
  }
  \subfigure[Goal, prediction, error]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_soesgp/fig_003_aie_experiment_plot_basic.pdf}
    \label{fig:basic_op_1000_soesgp_fig2}
  }
  \subfigure[Model sweep t = 500]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_soesgp/fig_002_colormeshmatrix_reduced.pdf}
    \label{fig:basic_op_1000_soesgp_fig3}
  }
  \subfigure[Model sweep t = 1000]{
    \includegraphics[width=0.47\textwidth]{img/basic_op_id_1000steps_soesgp/fig_004_colormeshmatrix_reduced.pdf}
    \label{fig:basic_op_1000_soesgp_fig4}
  }
  \end{center}
  \caption{\label{fig:basic_op_1000_soesgp}Basic operation 1000 timesteps GP.}

\end{figure}

\subsection{Basic behavioural example M2}
\label{sec:exprs_basic_ex_m2}
\subsection{Basic behavioural example E2P}
\label{sec:exprs_basic_ex_e2p}
